Downstream analysis: A pipeline to analyse time-course gene expression data

### 2019-04-30 scripts/03_clustering.Rmd
FIXes and more

# Only assign labels to X% of the genes
max_score = quantile(rowMin(all_scores), c(percentage_genes_to_label))
genes_to_not_consider = rowMin(all_scores) >= max_score
labels = apply(all_scores, 1, which.min)
labels[genes_to_not_consider] = NA
names(labels) = row.names(data)
```

```{r fig.height=8, fig.width=8}
par(mfrow=c(4, 5), mar=c(0.1, 0.1, 0.1, 0.1))
for(k in 1:n_clusters){
    hist(all_scores[k, ], breaks=20, col="black")
    abline(v=max_score, col="#AB0000")
}
```

@@

```{r}
table(kmeans_clusters$clusters)
barplot(table(kmeans_clusters$clusters), col="black")
```

```{r}
table(labels)
barplot(table(labels), col="black")
```

@@

```{r fig.width=8, fig.height=8}
gene_names = row.names(data)
genes_to_plot =  names(kmeans_clusters$clusters[kmeans_clusters$clusters == 15])
data_to_plot = data[genes_to_plot, ]
submeta = meta
submeta$Group = factor(submeta$Group, levels(submeta$Group)[c(3, 1, 2, 5, 4)])
ord = order(
  submeta$Group,
  submeta$Timepoint,
  submeta$Replicate)
submeta$Timepoint = as.factor(submeta$Timepoint)

@@

NMF::aheatmap(
    moanin:::rescale_values(data_to_plot),
    Colv=NA,
    annCol=submeta[,
                c("Group", "Timepoint")],
    annLegend=FALSE,
    annColors=ann_colors,
    main="Cluster 15",
    treeheight=0)

>>>>

all_scores = as.matrix(all_scores)
# Only assign labels to X% of the genes
max_score = quantile(rowMin(all_scores), c(percentage_genes_to_label))
genes_to_not_consider = rowMin(all_scores) >= max_score
labels = apply(all_scores, 1, which.min)
labels[genes_to_not_consider] = NA
names(labels) = row.names(data)
```

@@

```{r}
barplot(table(kmeans_clusters$clusters), col="black")
```

```{r}
barplot(table(labels), col="black")
labels = kmeans_clusters$clusters
```

@@

```{r fig.width=8, fig.height=8}
cluster_to_plot = 13
gene_names = row.names(data)
genes_to_plot =  names(kmeans_clusters$clusters[kmeans_clusters$clusters == cluster_to_plot])
data_to_plot = data[genes_to_plot, ]
submeta = meta
submeta$Group = factor(submeta$Group, levels(submeta$Group)[c(3, 1, 2, 5, 4)])
ord = order(
  submeta$Group,
  submeta$Timepoint,
  submeta$Replicate)
submeta$Timepoint = as.factor(submeta$Timepoint)

@@

NMF::aheatmap(
    moanin:::rescale_values(data_to_plot),
    Colv=NA,
    annCol=submeta[,
                c("Group", "Timepoint")],
    annLegend=FALSE,
    annColors=ann_colors,
    main=paste("Cluster", cluster_to_plot),
    treeheight=0)

### 2019-04-30 scripts/04_downstream_analysis.Rmd
FIXes and more

```{r enrichment_analysis, results="hide", message=FALSE}
clusters = sort(as.vector(unlist(unique(labels))))
clusters = c(1, 2, 3)
all_pathways = data.frame()
for(cluster in clusters){
    genes = gene_names[labels == cluster]

    # convert gene names
    genes = getBM(attributes=c("ensembl_gene_id", "entrezgene"),
		  filters="refseq_mrna", values=genes,
		  mart=ensembl)["entrezgene"]
    genes = as.vector(unlist(genes))
    pathways = KEGGprofile::find_enriched_pathway(
	genes, species="mmu",
	download_latest=TRUE)

    if((dim(all_pathways)[1] == 0) & (dim(pathways$stastic)[1] != 0)){
	print("Initialize")
	pathways$stastic["Cluster"] = cluster
	all_pathways = pathways$stastic
    }else if(dim(pathways$stastic)[1] != 0) {
	print("Merge")
	pathways$stastic["Cluster"] = cluster
	all_pathways = merge(all_pathways, pathways$stastic, all=TRUE)
    }
}
```

@@

all_pathways = all_pathways[
    order(all_pathways["Cluster"],
	  all_pathways["pvalueAdj"]),]

# Select the columns to print
all_pathways = subset(
    all_pathways,
    select=c(
	"Pathway_Name", "Percentage",
	"pvalueAdj", "Cluster"))

colnames(all_pathways) = c("Pathway", "Percentage", "Adj. p-value", "Cluster")
all_pathways["Percentage"] = all_pathways["Percentage"] * 100
kable(all_pathways, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

@@

cluster = c(1, 2, 3)
ontologies = c("BP", "NF", "CC")
# It's taking too long to compile with everything. Just limit ourselves for
# now to BP
ontologies = c("BP")
all_go_terms = data.frame()
for(cluster in clusters){
    for(ontology in ontologies){ 
	scores = as.numeric(labels == cluster)
	names(scores) = gene_names
	go_terms_enriched = find_enriched_go_terms(
	    scores,
	    gene_id_go_mapping, ontology=ontology)

	if((dim(all_go_terms)[1] == 0) & (dim(go_terms_enriched)[1] != 0)){
	    go_terms_enriched["Cluster"] = cluster
	    go_terms_enriched["Ontology"] = ontology
	    all_go_terms = go_terms_enriched
	}else if(dim(go_terms_enriched)[1] != 0){
	    go_terms_enriched["Cluster"] = cluster
	    go_terms_enriched["Ontology"] = ontology
	    all_go_terms = merge(all_go_terms, go_terms_enriched, all=TRUE)
	}
    }
}
```

@@

all_go_terms = all_go_terms[
    order(all_go_terms[, "Cluster"],
	  all_go_terms[, "Ontology"],
	  all_go_terms[, "resultFisher_padj"]), ]

kable(all_go_terms, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

>>>>

```{r enrichment_analysis, results="hide", message=FALSE}
cluster = 13
genes = gene_names[labels == cluster]

# convert gene names
genes = getBM(attributes=c("ensembl_gene_id", "entrezgene"),
		filters="refseq_mrna", values=genes,
		mart=ensembl)["entrezgene"]
genes = as.vector(unlist(genes))
pathways = KEGGprofile::find_enriched_pathway(
    genes, species="mmu",
    download_latest=TRUE)$stastic

```

@@

pathways = pathways[
    order(pathways["pvalueAdj"]),]

# Select the columns to print
pathways = subset(
    pathways,
    select=c(
	"Pathway_Name", "Percentage",
	"pvalueAdj"))

colnames(pathways) = c("Pathway", "Percentage", "Adj. p-value")
pathways["Percentage"] = pathways["Percentage"] * 100
kable(pathways, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

@@

scores = as.numeric(labels == cluster)
names(scores) = gene_names
go_terms_enriched = find_enriched_go_terms(
    scores,
    gene_id_go_mapping, ontology="BP")

```

@@

go_terms_enriched = go_terms_enriched[
    order(go_terms_enriched[, "resultFisher_padj"]), ]

kable(go_terms_enriched, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```
